---
title: "Publicly Available Voter Registration Data - What Can We Learn"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("ggplot2")
library("ggmap")
library("dplyr")
library("lubridate")

df <- read.csv("~/Desktop/Vote/votingdata_enhanced.csv")
df$lat3 <- round(df$lat,3)
df$lon3 <- round(df$lon,3)

```

## Registered Voters in Washington DC

The District of Columbia Board of Elections publicly released voter registration data as for 435,306 registered voters in Washington, DC in May 2016, which can be accessed at the following link <https://www.dcboee.org/pdf_files/ListOfRegisteredVoters051616.pdf>. Whether or not this data should be posted publically is a good question, but nothing should ever be posted as a 9,000 page .pdf file! Here's a quick look into what kind of aggregate statistics can we learn from this data after transforming it into a workable form. 

(1) There are 336,000 registered Democrats, 28,000 registered Republicans, and 72,000 individuals not registered with either party. That means 77% of registered voters are Democrats and only 6% are Republicans!
```{r, echo=FALSE, results="asis", warning=FALSE, message=FALSE}
# df %>% group_by(party) %>% summarize(count=n()) %>% mutate(count/sum(count))
```
(2) The highest density of registered Republicans is located on Boling Airforce Base, which contains the only two square kilometers with more than 50% registered Republicans.
```{r, echo=FALSE, results="asis", warning=FALSE, message=FALSE}
# df %>% group_by(party, round(lat,2), round(lon,2)) %>% summarize(count=n()) %>% mutate(p=count/sum(count)) %>% filter(party=="REP") %>% arrange(-p)
```
(3) We can map the data to see where the greatest density of Democrats, Republicans, and Party-less voters are. 
```{r, fig.width=9, fig.height=9, echo=FALSE, results="asis", warning=FALSE, message=FALSE}
data3 <- df %>% group_by(lat3,lon3, party) %>% summarize(count=n()) %>% group_by(lat3,lon3) %>% mutate(Proportion=count/sum(count)) %>% filter(party=="DEM")
loc <- c(lon = -77.02, lat =  38.9)
dc.map = get_map(location = loc, zoom = 11, color = "bw", maptype="watercolor")
ggmap(dc.map, extent = "normal", maprange=FALSE)  + geom_point(data=data3,aes(y=lat3,x=lon3, color=Proportion), size=.01) +
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Democrat Voters") +
  scale_colour_gradient(low="red",high="blue") +
  theme(plot.title =element_text(size=18), text=element_text(size=12),
        panel.background = element_rect(fill = 'white', colour = 'white'),
        legend.text=element_text(size=12),
        panel.grid = element_blank(),
        axis.text=element_text(size=10))

data3 <- df %>% group_by(lat3,lon3, party) %>% summarize(count=n()) %>% group_by(lat3,lon3) %>% mutate(Proportion=count/sum(count)) %>% filter(party=="REP")
ggmap(dc.map, extent = "normal", maprange=FALSE)  + geom_point(data=data3,aes(y=lat3,x=lon3, color=Proportion), size=.01) +
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Republican Voters") +
  scale_colour_gradient(low="blue",high="red") +
  theme(plot.title =element_text(size=18), text=element_text(size=12),
        panel.background = element_rect(fill = 'white', colour = 'white'),
        legend.text=element_text(size=12),
        panel.grid = element_blank(),
        axis.text=element_text(size=10))

data3 <- df %>% group_by(lat3,lon3, party) %>% summarize(count=n()) %>% group_by(lat3,lon3) %>% mutate(Proportion=count/sum(count)) %>% filter(party=="N-P")
ggmap(dc.map, extent = "normal", maprange=FALSE)  + geom_point(data=data3,aes(y=lat3,x=lon3, color=Proportion), size=.01) +
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Party-Less Voters") +
  scale_colour_gradient(low="green",high="black") +
  theme(plot.title =element_text(size=18), text=element_text(size=12),
        panel.background = element_rect(fill = 'white', colour = 'white'),
        legend.text=element_text(size=12),
        panel.grid = element_blank(),
        axis.text=element_text(size=10))
```
(4) 40% of the registered voters have registered since 2010, and over 10,000 voters registered before 1980. Since 2000, voter registration has peaked around election years.
```{r, echo=FALSE, results="asis", warning=FALSE, message=FALSE}
df$date <- mdy(df$registrationdate)
ggplot(df[df$date<"2016-11-08",],aes(x=date)) + geom_density() + 
  xlab("Date") +
  ylab("Proportion") +
  ggtitle("Distribution of Registration Dates") +  
  theme(plot.title =element_text(size=18), text=element_text(size=12), 
        panel.background = element_rect(fill = 'white', colour = 'white'),
        legend.text=element_text(size=12),
        panel.grid = element_blank(),
        axis.text=element_text(size=10))

# df$year <- year(df$date)
# View(df %>% group_by(year) %>% summarize(count=n()) %>% mutate(p=count/sum(count)))
```
(5) And ward 7 has the highest proportion of registered Democrats.
```{r kable, echo=FALSE, results="asis", warning=FALSE, message=FALSE}
data <- data.frame(df %>% group_by(ward, party) %>% summarize(count=n()) %>% group_by(ward) %>% mutate(p=round(count/sum(count),2)))
colnames(data) <- c("Ward","Party","Count","Proportion")
knitr::kable(data)
```